State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-06-01 StateOfPractice/Papers/SOP_MI_OSR/SOP_MI_OSR.tex
Edits up to and including the Results section

We grade the selected software using the measurement template summarized in
\citet{SmithEtAl2021}.  The template provides measures of the qualities listed
in Section~\ref{sec_software_quality}.
For each software package, we fill in the template questions. To stay within the
target of 160 person hours to measure the domain, we allocated between one and
four hours for each package. Project developers can be contacted for help
regarding installation, if necessary, but we impose a cap of about two hours on
the installation process, to keep the overall measurement time feasible.
Figure~\ref{fg_grading_template_example} shows an excerpt of the spreadsheet.
The spreadsheet includes a column for each measured software package. TODO - point to full data online.

@@

Saaty developed AHP in the 1970s, and people have widely used it since to make
and analyze multiple criteria decisions \citep{VaidyaEtAl2006}. AHP organizes
multiple criteria in a hierarchical structure and uses pairwise comparisons
between alternatives to calculate relative ratios \citep{Saaty1990}. AHP works
with sets of $n$ \textit{options} and $m$ \textit{criteria}.  In our project
$n=29$ and $m=9$ since there are 29 options (software products) and 9 criteria
(qualities). We rank the software for each of the qualities, and then we combine
the quality rankings into an overall ranking based on the relative priorities
between qualities.

@@

We initially identified 48 MI candidate software projects from the literature
\citep{Bjorn2017, Bruhschwein2019, Haak2015}, on-line articles \citep{Emms2019,
Hasan2020, Mu2019}, and forum discussions \citep{Samala2014}.  The full list of
48 packages is available in \citet{Dong2021}.  To reduce the length of the list
to a manageable number (29 in this case, as given in Section~\ref{ch_results}),
we filtered the original list as follows:

@@

\item We removed the packages that did not have source code available, such as
\textit{MicroDicom}, \textit{Aliza}, and \textit{jivex}.

@@

The Domain Expert provided a list of his top 12 software packages.  We compared
his list to our list of 29.  We found 6 packages were on both lists: \textit{3D
Slicer}, \textit{Horos}, \textit{ImageJ}, \textit{Fiji}, \textit{MRIcron} (we
actually use the update version \textit{MRIcroGL}) and \textit{Mango} (we
actually use the web version \textit{Papaya}).  Six software packages
(\textit{AFNI}, \textit{FSL}, \textit{Freesurfer}, \textit{Tarquin},
\textit{Diffusion Toolkit}, and \textit{MRItrix}) were on the Domain Expert
list, but not on our filtered list.  However, when we examined those packages,
we found they were out of scope, since their primary function was not
visualization.  The Domain Expert agreed with our final choice of 29 packages.

@@

Figure~\ref{fg_reliability_scores} shows the AHP results.  As shown in
Section~\ref{sec_result_installability}, most of the software products did not
``break'' during installation, or did not need installation; \textit{dwv} and
\textit{GATE} broke in the building stage, and the processes were not
recoverable; we could not install the dependency for \textit{DICOM Viewer}. Of
the seven software packages with a getting started tutorial and operation steps
in the tutorial, most showed no error when we followed the steps. However,
\textit{GATE} could not open macro files and became unresponsive several times,
without any descriptive error message. When assessing robustness
(Section~\ref{sec_result_robustness}), we found that \textit{Drishti} crashed
when loading damaged image files, without showing any descriptive error message.
We did not find any problems with the on-line version of
\textit{dwv}.

@@

Figure~\ref{fg_maintainability_scores} shows the ranking results for
maintainability. We gave \textit{3D Slicer} the highest score because we found
it to have the most comprehensive artifacts. For example, as far as we could
find, only a few of the 29 projects had a product, developer's manual, or API
documentation, and only \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}
included all three documents. Moreover, \textit{3D Slicer} has a much higher
percentage of closed issues (92\%) compared to \textit{ImageJ} (52\%) and
\textit{Fiji} (64\%). Table~\ref{tab_maintainability_docs} shows which
projects had these documents, in descending order of their maintainability
scores.

@@

Figure~\ref{fg_surface_understandability_scores} shows the scores for surface
understandability. All projects had a consistent coding style with parameters in
the same order for all functions; modularized code; and, clear comments, indicating
what is done, not how. However, we only found explicit identification of a
coding standard for 3 out of the 29: \textit{3D Slicer}, \textit{Weasis}, and
\textit{ImageJ}. We also found hard-coded constants (rather than symbolic
constants) in \textit{medInria}, \textit{dicompyler}, \textit{MicroView}, and
\textit{Papaya}. We did not find any reference to the algorithms used in
projects \textit{XMedCon}, \textit{DicomBrowser}, \textit{3DimViewer},
\textit{BioImage Suite Web}, \textit{Slice:Drop}, \textit{MatrixUser},
\textit{DICOM Viewer}, \textit{dicompyler}, and \textit{Papaya}.

>>>>

We grade the selected software using the measurement template summarized in
\citet{SmithEtAl2021}.  The template provides measures of the qualities listed
in Section~\ref{sec_software_quality}.
For each software package, we fill in the template questions. To stay within the
target of 160 person hours to measure the domain, we allocated between one and
four hours for each package. Project developers can be contacted for help
regarding installation, if necessary, but we impose a cap of about two hours on
the installation process, to keep the overall measurement time feasible.
Figure~\ref{fg_grading_template_example} shows an excerpt of the spreadsheet.
The spreadsheet includes a column for each measured software package. \citep{Dong2021-Data} provides the full set of measurement data.

@@

Saaty developed AHP in the 1970s, and people have widely used it since to make
and analyze multiple criteria decisions \citep{VaidyaEtAl2006}. AHP organizes
multiple criteria in a hierarchical structure and uses pairwise comparisons
between alternatives to calculate relative ratios \citep{Saaty1990}. AHP works
with sets of $n$ \textit{options} and $m$ \textit{criteria}.  In our project
$n=29$ and $m=9$ since there are 29 options (software products) and 9 criteria
(qualities). With AHP the sum of the grades (scores) for all products for a
given quality will be 1.0.  We rank the software for each of the qualities, and
then we combine the quality rankings into an overall ranking based on the
relative priorities between qualities.

@@

We initially identified 48 MI candidate software projects from the literature
\citep{Bjorn2017, Bruhschwein2019, Haak2015}, on-line articles \citep{Emms2019,
Hasan2020, Mu2019}, and forum discussions \citep{Samala2014}.  The full list of
48 packages is available in \citet{Dong2021}.  To reduce the length of the list
to a manageable number (29 in this case), we filtered the original list as
follows:

@@

\item We removed the packages with no source code available, such as
\textit{MicroDicom}, \textit{Aliza}, and \textit{jivex}.

@@

The Domain Expert provided a list of his top 12 software packages.  We compared
his list to our list of 29.  We found 6 packages were on both lists: \textit{3D
Slicer}, \textit{Horos}, \textit{ImageJ}, \textit{Fiji}, \textit{MRIcron} (we
actually use the update version \textit{MRIcroGL}) and \textit{Mango} (we
actually use the web version \textit{Papaya}).  Six software packages
(\textit{AFNI}, \textit{FSL}, \textit{Freesurfer}, \textit{Tarquin},
\textit{Diffusion Toolkit}, and \textit{MRItrix}) were on the Domain Expert
list, but not on our filtered list.  When we examined those packages,
we found they were out of scope, since their primary function was not
visualization.  The Domain Expert agreed with our final choice of 29 packages.

@@

Figure~\ref{fg_reliability_scores} shows the AHP results.  As shown in
Section~\ref{sec_result_installability}, most of the software products did not
``break'' during installation, or did not need installation; \textit{dwv} and
\textit{GATE} broke in the building stage, and the processes were not
recoverable; we could not install the dependency for \textit{DICOM Viewer}. Of
the seven software packages with a getting started tutorial and operation steps
in the tutorial, most showed no error when we followed the steps. However,
\textit{GATE} could not open macro files and became unresponsive several times,
without any descriptive error message. We found that \textit{Drishti} crashed
when loading damaged image files, without showing any descriptive error message.
We did not find any problems with the on-line version of \textit{dwv}.

@@

Figure~\ref{fg_maintainability_scores} shows the ranking results for
maintainability. We gave \textit{3D Slicer} the highest score because we found
it to have the most comprehensive artifacts. Only a few of the 29 projects had a
product, developer's manual, or API documentation, and only \textit{3D Slicer},
\textit{ImageJ}, \textit{Fiji} included all three documents. Moreover,
\textit{3D Slicer} has a much higher percentage of closed issues (92\%) compared
to \textit{ImageJ} (52\%) and \textit{Fiji} (64\%).
Table~\ref{tab_maintainability_docs} shows which projects had these documents,
in descending order of their maintainability scores.

@@

Figure~\ref{fg_surface_understandability_scores} shows the scores for surface
understandability. All projects had a consistent coding style with parameters in
the same order for all functions, modularized code, and, clear comments that
indicate what is done, not how. However, we only found explicit identification
of a coding standard for 3 out of the 29: \textit{3D Slicer}, \textit{Weasis},
and \textit{ImageJ}. We also found hard-coded constants (rather than symbolic
constants) in \textit{medInria}, \textit{dicompyler}, \textit{MicroView}, and
\textit{Papaya}. We did not find any reference to the algorithms used in
projects \textit{XMedCon}, \textit{DicomBrowser}, \textit{3DimViewer},
\textit{BioImage Suite Web}, \textit{Slice:Drop}, \textit{MatrixUser},
\textit{DICOM Viewer}, \textit{dicompyler}, and \textit{Papaya}.
